---
title: "Spatial questions - preliminary EDA"
output: html_document
---
This file contains the exploratory data analysis of the spatial questions' responses, using the `prelim_singlereview.csv` file which contains results from 268 papers. Spatial questions are Q8-Q11.


```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(igraph)
library(plotly)
library(gridExtra)
```


```{r read data, echo=FALSE}
non_excl_titles = read.csv("../data/cleaned/prelim_singlereview.csv") %>% 
  filter(qnum =='Q3', !answer=='Yes') %>%
  select(Title)
spat = read.csv("../data/cleaned/prelim_singlereview.csv") %>% 
  filter(Title %in% non_excl_titles$Title) %>%
  filter(qnum %in% c("Q8","Q9","Q10","Q11"))
```

## Q8: Does the paper consider or compare multiple spatial scales?

```{r q8, echo=FALSE}
spat %>% 
  filter(qnum == "Q8") %>%
  group_by(fullquestion) %>%
  summarise(Yes = sum(answer=='Yes', na.rm = T),
            No = sum(answer=='No', na.rm = T),
            'NA' = sum(is.na(answer))) %>%
  melt() %>%
  ggplot(aes(x = variable, y = value)) +
  geom_col() +
  geom_label(aes(label = value)) +
  ylab("Number of papers") +
  xlab("Q8: Does the paper consider or compare multiple spatial scales?") +
  theme_bw()
```


## Q9: What is the size of the unit of replication? Select multiple if inferences are made at multip...

### Number of papers at each scale

#### Plots
```{r plot q9, echo=FALSE}
p1 = spat %>% 
  filter(qnum == "Q9", abbr == "Plots", !is.na(answer), !answer==0) %>% 
  group_by(fullquestion, Group, answer) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(Group) %>%
  mutate(Group = factor(Group, levels = c('25m','50m','100m','500m','1km','10km', '100km','1000km','100Mgm','100Gm','101Gm','unk'))) %>%
  ggplot() +
  geom_col(aes(x = Group, y = count, group = answer, fill = answer)) +
  labs(x = "Size of plots", y = "Number of papers", fill = "Number of plots") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))

p2 = spat %>% 
  filter(qnum == "Q9", abbr == "Plots", !is.na(answer), !answer==0) %>% 
  group_by(fullquestion, Group, answer) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(Group) %>%
  mutate(Group = factor(Group, levels = c('25m','50m','100m','500m','1km','10km', '100km','1000km','100Mgm','100Gm','101Gm','unk'))) %>%
  ggplot() +
  geom_col(aes(x = answer, y = count)) +
  labs(x = "Number of plots", y = "Number of papers") +
  theme_bw() 

grid.arrange(p1, p2, ncol = 2, widths = c(3,1))
```

#### Sites
```{r site q9, echo=FALSE}
p1 = spat %>% 
  filter(qnum == "Q9", abbr == "Sites", !is.na(answer), !answer==0) %>% 
  group_by(fullquestion, Group, answer) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(Group) %>%
  mutate(Group = factor(Group, levels = c('25m','50m','100m','500m','1km','10km', '100km','1000km','100Mgm','100Gm','101Gm','unk'))) %>%
  ggplot() +
  geom_col(aes(x = Group, y = count, group = answer, fill = answer)) +
  labs(x = "Size of sites", y = "Number of papers", fill = "Number of sites") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))

p2 = spat %>% 
  filter(qnum == "Q9", abbr == "Sites", !is.na(answer), !answer==0) %>% 
  group_by(fullquestion, Group, answer) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(Group) %>%
  mutate(Group = factor(Group, levels = c('25m','50m','100m','500m','1km','10km', '100km','1000km','100Mgm','100Gm','101Gm','unk'))) %>%
  ggplot() +
  geom_col(aes(x = answer, y = count)) +
  labs(x = "Number of sites", y = "Number of papers") +
  theme_bw()

grid.arrange(p1,p2,ncol=2,widths=c(3,1))
```


### Heatmap linking plots to sites at each scale
After looking at this, I'm not sure it's super useful, but it could be useful if we bin them into larger scale bins (like small, medium, large). With the right binning, it could potentially fit well into the scale figure.

```{r make network, echo=FALSE}
plot_site = spat %>% 
  filter(qnum=="Q9", abbr == "Plots", !is.na(answer), !answer==0) %>% 
  select(Title, answer, Group) %>%
  rename(plot_answer = answer, plot_group = Group) %>%
  full_join(spat  %>% 
              filter(qnum=="Q9", abbr == "Sites", !is.na(answer), !answer==0) %>% 
              select(Title, answer, Group) %>%
              rename(site_answer = answer, site_group = Group),
            by = 'Title') 


#node list
plots = spat %>%
  filter(qnum=="Q9", abbr=='Plots') %>%
  distinct(Group) %>%
  rename(label = Group)

sites = spat %>% 
  filter(qnum=="Q9", abbr=='Sites') %>%
  distinct(Group) %>%
  rename(label = Group)


nodes = plots %>% 
  full_join(sites, by = 'label') 
nodes$label = as.character(nodes$label)
nodes = nodes %>% 
  rbind(list("none")) %>% 
  rowid_to_column(var = "id")

plot_site$plot_group = plot_site$plot_group %>% as.character()
plot_site$site_group = plot_site$site_group %>% as.character()

edges = plot_site %>%
  replace_na(replace = list("plot_answer" = NA, "plot_group" = "none", "site_answer" = NA, "site_group" = "none")) %>% 
  group_by(plot_group, site_group) %>%
  summarize(weight = n()) %>% 
  ungroup() %>%
  left_join(nodes, by = c('plot_group'='label')) %>%
  rename(from = id) %>%
  left_join(nodes, by = c('site_group'='label')) %>%
  rename(to = id) %>%
  select(from, to, weight) %>%
  replace_na(replace = list("from" = nodes %>% 
                              filter(label=='none') %>% 
                              select(id) %>% 
                              as.numeric(), 
                            "to" = nodes %>% 
                              filter(label=='none') %>% 
                              select(id) %>% 
                              as.numeric()))

net = graph.data.frame(d = edges, vertices = nodes, directed = F)
# this could work without doing all of the site and plot specific stuff
netm <- get.adjacency(net, attr="weight", sparse=F) 
colnames(netm) <- V(net)$label
rownames(netm) <- V(net)$label
palf <- colorRampPalette(c("gold", "red")) 

c_sums = colSums(netm)
r_sums = rowSums(netm)
netm = rbind(netm, 'Total' = c_sums)
netm = cbind(netm, 'Total' = c(r_sums, NA)) #makes Total,Total entry 0 so it doesn't overwhelm the heatmap
```


This is a `plotly` plot, so you can click and hover to see the values.
```{r heatmap plotly, echo=FALSE, warning=FALSE}
p = ggplot(data=melt(netm[-nrow(netm),-ncol(netm)])) + #[-nrow(netm),-ncol(netm)] # to exclude the totals
  geom_raster(aes(x = Var1, y = Var2, fill = value, text = value)) +
  xlab('Plot size') +
  ylab('Site size') +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(fill = 'Number of papers') 

ggplotly(p, tooltip = 'text')
```

This plot shows the number of papers with plot size X and site size Y, which could help us see trends in the scales that studies are linking from their plots to their sites (but the definitions of sites and plots was a little tricky in the coding so this might take more thought).

### Also a heatmap for just those papers that examined patterns across scales

```{r make network cross scale, echo=FALSE}
crossscale_titles = spat %>% filter(qnum == "Q8") %>% filter(answer=='Yes') %>% select(Title)

plot_site = spat %>% 
  filter(qnum=="Q9", abbr == "Plots", !is.na(answer), !answer==0, Title %in% crossscale_titles$Title) %>% 
  select(Title, answer, Group) %>%
  rename(plot_answer = answer, plot_group = Group) %>%
  full_join(spat  %>% 
              filter(qnum=="Q9", abbr == "Sites", !is.na(answer), !answer==0) %>% 
              select(Title, answer, Group) %>%
              rename(site_answer = answer, site_group = Group),
            by = 'Title') 


#node list
plots = spat %>%
  filter(qnum=="Q9", abbr=='Plots') %>%
  distinct(Group) %>%
  rename(label = Group)

sites = spat %>% 
  filter(qnum=="Q9", abbr=='Sites') %>%
  distinct(Group) %>%
  rename(label = Group)


nodes = plots %>% 
  full_join(sites, by = 'label') 
nodes$label = as.character(nodes$label)
nodes = nodes %>% 
  rbind(list("none")) %>% 
  rowid_to_column(var = "id")

plot_site$plot_group = plot_site$plot_group %>% as.character()
plot_site$site_group = plot_site$site_group %>% as.character()

edges = plot_site %>%
  replace_na(replace = list("plot_answer" = NA, "plot_group" = "none", "site_answer" = NA, "site_group" = "none")) %>% 
  group_by(plot_group, site_group) %>%
  summarize(weight = n()) %>% 
  ungroup() %>%
  left_join(nodes, by = c('plot_group'='label')) %>%
  rename(from = id) %>%
  left_join(nodes, by = c('site_group'='label')) %>%
  rename(to = id) %>%
  select(from, to, weight) %>%
  replace_na(replace = list("from" = nodes %>% 
                              filter(label=='none') %>% 
                              select(id) %>% 
                              as.numeric(), 
                            "to" = nodes %>% 
                              filter(label=='none') %>% 
                              select(id) %>% 
                              as.numeric()))

net = graph.data.frame(d = edges, vertices = nodes, directed = F)
# this could work without doing all of the site and plot specific stuff
netm <- get.adjacency(net, attr="weight", sparse=F) 
colnames(netm) <- V(net)$label
rownames(netm) <- V(net)$label
palf <- colorRampPalette(c("gold", "red")) 

c_sums = colSums(netm)
r_sums = rowSums(netm)
netm = rbind(netm, 'Total' = c_sums)
netm = cbind(netm, 'Total' = c(r_sums, NA)) #makes Total,Total entry 0 so it doesn't overwhelm the heatmap
```


This is a `plotly` plot, so you can click and hover to see the values.
```{r heatmap plotly cross scale, echo=FALSE, warning=FALSE}
p = ggplot(data=melt(netm[-nrow(netm),-ncol(netm)])) + #[-nrow(netm),-ncol(netm)] # to exclude the totals
  geom_raster(aes(x = Var1, y = Var2, fill = value, text = value)) +
  xlab('Plot size') +
  ylab('Site size') +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(fill = 'Number of papers', title = "Given that the paper looked across scales") 

ggplotly(p, tooltip = 'text')
```







## Q10: Does the paper consider connectivity between spatial replicates (sites, plots)?

```{r connectivity, echo=FALSE}
spat %>%
  filter(qnum=='Q10') %>%
  group_by(fullquestion) %>%
  summarise(Yes = sum(answer=='Yes', na.rm = T),
            No = sum(answer=='No', na.rm = T),
            'NA' = sum(is.na(answer))) %>%
  melt() %>%
  ggplot(aes(x = variable, y = value)) +
  geom_col() +
  geom_label(aes(label = value)) +
  ylab("Number of papers") +
  xlab("Q10: Does the paper consider connectivity between spatial replicates (sites, plots)?") +
  theme_bw()
```

#### If paper examined connectivity...what scale?
```{r scale for connectivity papers, echo = FALSE}
connect_titles = spat %>%
  filter(qnum=='Q10', answer=='Yes') %>%
  select(Title)

#plots with connectivity
spat %>% 
  filter(qnum == "Q9", abbr == "Plots", !is.na(answer), !answer==0, Title %in% connect_titles$Title) %>% 
  group_by(fullquestion, Group, answer) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(Group) %>%
  mutate(Group = factor(Group, levels = c('25m','50m','100m','500m','1km','10km', '100km','1000km','100Mgm','100Gm','101Gm','unk'))) %>%
  ggplot() +
  geom_col(aes(x = Group, y = count, group = answer, fill = answer)) +
  labs(x = "Size of plots", y = "Number of papers", fill = "Number of plots") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))

#sites with connectivity
spat %>% 
  filter(qnum == "Q9", abbr == "Sites", !is.na(answer), !answer==0, Title %in% connect_titles$Title) %>% 
  group_by(fullquestion, Group, answer) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(Group) %>%
  mutate(Group = factor(Group, levels = c('25m','50m','100m','500m','1km','10km', '100km','1000km','100Mgm','100Gm','101Gm','unk'))) %>%
  ggplot() +
  geom_col(aes(x = Group, y = count, group = answer, fill = answer)) +
  labs(x = "Size of sites", y = "Number of papers", fill = "Number of sites") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))

```


## Q11: What is the distance between replicates? (Given that the paper measured connectivity)

```{r replicate distance, echo=FALSE}
#have to deal with the ability to check multiple boxes
answers = spat %>%
  filter(qnum=='Q11', Title %in% connect_titles$Title) %>%
  group_by(answer) %>%
  summarise(count = n()) %>% 
  select(answer) 

# collapse and separate to get all unique possibilities
vect_ans = paste(answers$answer, collapse= ',') 
uniq_answers = unlist(strsplit(vect_ans, split=',')) %>% unique()
count_df = spat %>%
  filter(qnum=='Q11', Title %in% connect_titles$Title) %>%
  group_by(answer) %>%
  summarise(count = n())

# grep (search) for each unique answer in the count_df, return the sum of the
# counts where that value was selected
ans_df = data.frame('answer' = uniq_answers, 'count' = NA)
for (i in 1:nrow(ans_df)) {
  rows_found = grep(as.character(ans_df$answer[i]), count_df$answer)
  ans_df$count[i] = sum(count_df$count[rows_found])
}

ans_df %>%
  ggplot() +
  geom_col(aes(x = answer, y = count)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0)) +
  xlab('Distance between replicates (unordered)') +
  ylab('Number of papers')

```

