---
title: "Spatial binning EDA"
output: html_document
---
```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(reshape2)
library(plotly)
```

Data used:
```{r read data, echo=FALSE}
data_path = "../data/cleaned/prelim_singlereview.csv"
df = read.csv(data_path)
non_excl_titles = df %>% 
  filter(qnum =='Q3', !answer=='Yes') %>%
  select(Title)
spat = df %>% 
  filter(Title %in% non_excl_titles$Title) %>%
  filter(qnum %in% c("Q8","Q9","Q10","Q11"))

print(data_path)
```

### Summary
From the preliminary data:

 * Lots of studies with 'unk' spatial scale (this means that no plot nor site scales were reported by coders). 
 * Lots of small scale studies, some large scale studies
 * Relatively few medium scale, and medium-large scale papers
 * Not that many papers cross the scale bins we've used
 
My takeaway (Grant): A fair amount of small scale studies reach up into the medium scale category, but hardly any large scale studies have smaller scale components. Way too many studies do not report (or we couldn't find) the spatial scale at which they are operating...maybe a lot of these are simulations? - doesn't look like it, see below


# Spatial binning to fit scale figure

The end result of these numbers may just be displayed as n=X in the existing scale figure, but there are other potential ways to display these things, too. 

Key for binning:
```{r, echo = FALSE}
smalls = c('25m','50m','100m','500m', '1km')
mediums = c('10km', '100km', '1000km')
larges = c('100Mgm','100Gm','101Gm')

scale_fig_df = spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group)) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  separate_rows(scales, sep = ",") %>% 
  mutate(scales_binned = case_when(
    scales %in% smalls ~ 'Small',
    scales %in% mediums ~ 'Medium',
    scales %in% larges ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>%
  group_by(Title) %>%
  summarise(cross_grps = paste0(unique(scales_binned), collapse = ',')) %>% 
  ungroup() %>%
  group_by(cross_grps) %>%
  summarise(count = n()) %>%
  mutate(Scale_groups = gsub("unk,","",cross_grps)) %>%
  mutate(Scale_groups = gsub(",unk","",Scale_groups)) %>%
  group_by(Scale_groups) %>%
  summarise(count = sum(count)) %>%
  mutate(Scale_groups = factor(Scale_groups, levels = c("Small", "Small,Medium","Medium","Medium,Large","Large","Small,Large","Small,Medium,Large","unk")))


key_df = data.frame("Scale_binned" = c('Small','Medium', 'Large'), "scales" = c(paste0(smalls,collapse = ","), paste0(mediums,collapse = ","), paste0(larges,collapse = ",")))
print(key_df)
```

Number of papers in each bin:
```{r, echo=FALSE}
print(arrange(scale_fig_df, Scale_groups))
# These numbers are not double counting papers. Each paper gets put into exactly one bin depending on which scales it examines. Other plots could include the total number of papers within each scale - if we ignore the between scale patterns it might make sense to double count some papers if what we care about is the individual scale.
```


```{r, echo = FALSE}
ggplot(data = scale_fig_df, aes(x = Scale_groups, y = count, label = count)) +
  geom_col() +
  geom_label() +
  xlab('Scale of study') +
  ylab('Number of papers') +
  theme_bw()
```


```{r, echo=FALSE, warning=FALSE}
cross_sc_tab = spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group)) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  separate_rows(scales, sep = ",") %>% 
  mutate(scales_binned = case_when(
    scales %in% c('25m','50m','100m','500m', '1km') ~ 'Small',
    scales %in% c('10km', '100km', '1000km') ~ 'Medium',
    scales %in% c('100Mgm','100Gm','101Gm') ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>%
  group_by(Title) %>%
  summarise(cross_grps = paste0(unique(scales_binned), collapse = ',')) %>% 
  ungroup() %>%
  group_by(cross_grps) %>%
  summarise(count = n()) %>%
  mutate(cross_grps_nounk = gsub("unk,","",cross_grps)) %>%
  mutate(cross_grps_nounk = gsub(",unk","",cross_grps_nounk)) %>%
  group_by(cross_grps_nounk) %>%
  summarise(count = sum(count)) %>%
  mutate(cross_grps_nounk = factor(cross_grps_nounk, levels = c("Small", "Small,Medium","Medium","Medium,Large","Large","Small,Large","Small,Medium,Large","unk"))) %>%
  filter(!cross_grps_nounk %in% c("Small,Medium,Large","unk")) %>%
  separate(cross_grps_nounk, into = c('scale1','scale2'), sep = ",") %>%
  mutate(scale2 = ifelse(is.na(scale2),scale1,scale2))
  

# matrix for heatmap
mat = matrix(nrow = length(unique(cross_sc_tab$scale1)), ncol = length(unique(cross_sc_tab$scale1)), dimnames = list(unique(cross_sc_tab$scale1),unique(cross_sc_tab$scale2)))

for (i in 1:nrow(mat)) {
  r_ = unique(cross_sc_tab$scale1)[i]
  for(j in 1:i){
    c_ = unique(cross_sc_tab$scale2)[j]
    mat[r_,c_] = cross_sc_tab %>% filter(scale1 == r_, scale2 == c_) %>% pull(count)
  }
  
}

p1 = 
  t(mat) %>% 
  melt() %>%
  mutate(Var1 = factor(Var1, levels = c("Small", "Medium","Large")),
         Var2 = factor(Var2, levels = c("Small", "Medium","Large")),
         label_ = ifelse(Var1==Var2, paste0('only ', Var1), NA)) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value, label = label_)) +
  geom_raster() +
  geom_label() + # doesn't work with plotly
  theme(axis.text.x = element_text(angle = 45)) +
  labs(fill = "Number of papers", x = "", y = "", title = "Scales of study") +
  theme_bw() 
# the numbers from this plot should be able to directly fill in the numbers from the scale plot
# as long as we remember to add the unknown scale and s-m-l numbers, since they don't fit on this plot


ggplotly(p1)
```


### Unknown scale studies...

```{r, echo=FALSE, warning=FALSE}
methods_used_df = df %>%
  filter(qnum=="Q6", !abbr %in% c('GenInfo','MethodsNotes')) %>%
  separate_rows(answer, sep = ",") %>%
  mutate(answer = ifelse(answer=="Observational (Includes data observed in situ OR via remote sensing", "Observational", as.character(answer))) %>%
  mutate(answer = ifelse(answer==" if used directly)", NA, as.character(answer))) %>%
  group_by(Title) %>%
  summarise(methods_used = paste0(unique(answer), collapse = ","))


spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group)) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  left_join(methods_used_df, by = "Title") %>%
  separate_rows(scales, sep = ",") %>%
  mutate(methods_used = gsub(",NA","",methods_used)) %>%
  mutate(scales_binned = case_when(
    scales %in% smalls ~ 'Small',
    scales %in% mediums ~ 'Medium',
    scales %in% larges ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>% 
  group_by(Title) %>%
  mutate(scales_binned = paste0(scales_binned, collapse = ",")) %>%
  filter(scales_binned == 'unk') %>%
  group_by(methods_used) %>%
  summarise(unk_count = n()) %>%
  mutate(methods_used = fct_reorder(methods_used, -unk_count)) %>%
  ggplot() +
  geom_col(aes(x = methods_used, y = unk_count)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +
  ylab("Number of papers") +
  ggtitle("For papers with unknown spatial scale, what type?")
```

Unclear what explains why so many studies had unknown spatial scale...

### Small scale studies...

```{r, echo=FALSE, warning=FALSE}
spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group)) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  left_join(methods_used_df, by = "Title") %>%
  separate_rows(scales, sep = ",") %>%
  mutate(methods_used = gsub(",NA","",methods_used)) %>%
  mutate(scales_binned = case_when(
    scales %in% smalls ~ 'Small',
    scales %in% mediums ~ 'Medium',
    scales %in% larges ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>% 
  group_by(Title) %>%
  mutate(scales_binned = paste0(scales_binned, collapse = ",")) %>%
  filter(scales_binned == 'Small') %>%
  group_by(methods_used) %>%
  summarise(small_count = n()) %>%
  mutate(methods_used = fct_reorder(methods_used, -small_count)) %>%
  ggplot() +
  geom_col(aes(x = methods_used, y = small_count)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +
  ylab("Number of papers") +
  ggtitle("For papers with small spatial scale, what type?")
```

### And large scale studies

```{r, echo=FALSE, warning=FALSE}
spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group)) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  left_join(methods_used_df, by = "Title") %>%
  separate_rows(scales, sep = ",") %>%
  mutate(methods_used = gsub(",NA","",methods_used)) %>%
  mutate(scales_binned = case_when(
    scales %in% smalls ~ 'Small',
    scales %in% mediums ~ 'Medium',
    scales %in% larges ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>% 
  group_by(Title) %>%
  mutate(scales_binned = paste0(scales_binned, collapse = ",")) %>%
  filter(scales_binned == 'Large') %>%
  group_by(methods_used) %>%
  summarise(small_count = n()) %>%
  mutate(methods_used = fct_reorder(methods_used, -small_count)) %>%
  ggplot() +
  geom_col(aes(x = methods_used, y = small_count)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") +
  ylab("Number of papers") +
  ggtitle("For papers with large spatial scale, what type?")
```

## Just for those studies with a 'Yes' for *Q8: Does the paper consider or compare multiple spatial scales?*
```{r, echo=FALSE}
crossscale_titles = spat %>% filter(qnum == "Q8") %>% filter(answer=='Yes') %>% pull(Title)

smalls = c('25m','50m','100m','500m', '1km')
mediums = c('10km', '100km', '1000km')
larges = c('100Mgm','100Gm','101Gm')

scale_fig_df_cross_sc = spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group), Title %in% crossscale_titles) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  separate_rows(scales, sep = ",") %>% 
  mutate(scales_binned = case_when(
    scales %in% smalls ~ 'Small',
    scales %in% mediums ~ 'Medium',
    scales %in% larges ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>%
  group_by(Title) %>%
  summarise(cross_grps = paste0(unique(scales_binned), collapse = ',')) %>% 
  ungroup() %>%
  group_by(cross_grps) %>%
  summarise(count = n()) %>%
  mutate(Scale_groups = gsub("unk,","",cross_grps)) %>%
  mutate(Scale_groups = gsub(",unk","",Scale_groups)) %>%
  group_by(Scale_groups) %>%
  summarise(count = sum(count)) %>%
  mutate(Scale_groups = factor(Scale_groups, levels = c("Small", "Small,Medium","Medium","Medium,Large","Large","Small,Large","Small,Medium,Large","unk")))


key_df = data.frame("Scale_binned" = c('Small','Medium', 'Large'), "scales" = c(paste0(smalls,collapse = ","), paste0(mediums,collapse = ","), paste0(larges,collapse = ",")))
print(key_df)

```

Number of papers in each bin:
```{r, echo=FALSE}
print(arrange(scale_fig_df_cross_sc, Scale_groups))
# These numbers are not double counting papers. Each paper gets put into exactly one bin depending on which scales it examines. Other plots could include the total number of papers within each scale - if we ignore the between scale patterns it might make sense to double count some papers if what we care about is the individual scale.
```


```{r, echo = FALSE}
ggplot(data = scale_fig_df_cross_sc, aes(x = Scale_groups, y = count, label = count)) +
  geom_col() +
  geom_label() +
  xlab('Scale of study') +
  ylab('Number of papers') +
  theme_bw()
```

```{r, echo=FALSE, warning=FALSE}
cross_sc_tab = spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group), Title %in% crossscale_titles) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  separate_rows(scales, sep = ",") %>% 
  mutate(scales_binned = case_when(
    scales %in% c('25m','50m','100m','500m', '1km') ~ 'Small',
    scales %in% c('10km', '100km', '1000km') ~ 'Medium',
    scales %in% c('100Mgm','100Gm','101Gm') ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>%
  group_by(Title) %>%
  summarise(cross_grps = paste0(unique(scales_binned), collapse = ',')) %>% 
  ungroup() %>%
  group_by(cross_grps) %>%
  summarise(count = n()) %>%
  mutate(cross_grps_nounk = gsub("unk,","",cross_grps)) %>%
  mutate(cross_grps_nounk = gsub(",unk","",cross_grps_nounk)) %>%
  group_by(cross_grps_nounk) %>%
  summarise(count = sum(count)) %>%
  mutate(cross_grps_nounk = factor(cross_grps_nounk, levels = c("Small", "Small,Medium","Medium","Medium,Large","Large","Small,Large","Small,Medium,Large","unk"))) %>%
  filter(!cross_grps_nounk %in% c("Small,Medium,Large","unk")) %>%
  separate(cross_grps_nounk, into = c('scale1','scale2'), sep = ",") %>%
  mutate(scale2 = ifelse(is.na(scale2),scale1,scale2))
  

# matrix for heatmap
mat = matrix(nrow = length(unique(cross_sc_tab$scale1)), ncol = length(unique(cross_sc_tab$scale1)), dimnames = list(unique(cross_sc_tab$scale1),unique(cross_sc_tab$scale2)))

for (i in 1:nrow(mat)) {
  r_ = unique(cross_sc_tab$scale1)[i]
  for(j in 1:i){
    c_ = unique(cross_sc_tab$scale2)[j]
    mat[r_,c_] = cross_sc_tab %>% filter(scale1 == r_, scale2 == c_) %>% pull(count)
  }
  
}

p1 = 
  t(mat) %>% 
  melt() %>%
  mutate(Var1 = factor(Var1, levels = c("Small", "Medium","Large")),
         Var2 = factor(Var2, levels = c("Small", "Medium","Large")),
         label_ = ifelse(Var1==Var2, paste0('only ', Var1), NA)) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value, label = label_)) +
  geom_raster() +
  geom_label() + # doesn't work with plotly
  theme(axis.text.x = element_text(angle = 45)) +
  labs(fill = "Number of papers", x = "", y = "", title = "Scales of study") +
  theme_bw() 
# the numbers from this plot should be able to directly fill in the numbers from the scale plot
# as long as we remember to add the unknown scale and s-m-l numbers, since they don't fit on this plot


ggplotly(p1)
```


# Now let's look at the combination of study type, spatial scale, and temporal scale
These plots have some tough-to-look-at colors (for me at least), and we might need to think about how to break up the groups, but I think they're helpful nonetheless.

Labels indicate total counts.

```{r, echo=FALSE}
time_trends_titles = df %>%
  filter(qnum == "Q7", abbr=='TimeTrends', answer=="Yes (e.g., a site sampled multiple points in time, a model with dynamics)") %>%
  pull(Title)

numyears_df = df %>% 
  filter(qnum == "Q7", abbr == 'YrsData', Title %in% time_trends_titles) %>%
  select(Title, answer) %>%
  rename(num_years = answer)

methods_used_df = df %>%
  filter(qnum=="Q6", !abbr %in% c('GenInfo','MethodsNotes')) %>%
  separate_rows(answer, sep = ",") %>%
  mutate(answer = ifelse(answer=="Observational (Includes data observed in situ OR via remote sensing", "Observational", as.character(answer))) %>%
  mutate(answer = ifelse(answer==" if used directly)", NA, as.character(answer))) %>%
  group_by(Title) %>%
  summarise(methods_used = paste0(unique(answer), collapse = ","))


spat_temp_methods_df = spat %>% 
  filter(qnum == 'Q9', !is.na(answer), !answer==0, !is.na(Group)) %>% 
  group_by(Title) %>%
  summarise(scales = paste0(unique(Group), collapse = ',')) %>%
  separate_rows(scales, sep = ",") %>% 
  mutate(scales_binned = case_when(
    scales %in% smalls ~ 'Small',
    scales %in% mediums ~ 'Medium',
    scales %in% larges ~ 'Large',
    scales == 'unk' ~ 'unk'
  )) %>%
  group_by(Title) %>%
  summarise(cross_grps = paste0(unique(scales_binned), collapse = ',')) %>% 
  ungroup() %>% 
  mutate(cross_grps = gsub("unk,","",cross_grps)) %>%
  mutate(cross_grps = gsub(",unk","",cross_grps)) %>%
  left_join(numyears_df, by = 'Title') %>%
  left_join(methods_used_df, by = 'Title') %>%
  rename(spat_scale = cross_grps) %>%
  mutate(methods_used = gsub(",NA","",methods_used)) %>%
  mutate(spat_scale = gsub("unk", NA, spat_scale)) %>%
  mutate(num_years = factor(num_years, levels = c('1 year or less','2–5 years','6–10 years','10+ years'))) %>%
  mutate(spat_scale = factor(spat_scale, levels = c("Small", "Small,Medium","Medium","Medium,Large","Large","Small,Large","Small,Medium,Large","unk")))

```

#### For each temporal scale, what is the spatial scale?
```{r, echo=FALSE, warning=FALSE}
spat_temp_methods_df %>% 
  group_by(methods_used,num_years,spat_scale) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = num_years, y = count)) +
  geom_bar(position = 'fill', stat='identity', aes(fill = spat_scale)) +
  geom_label(data = spat_temp_methods_df %>% group_by(num_years) %>% summarise(count = n()), 
             aes(label = count, y = 1.05)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) +
  ylab('Proportion of studies')

  #long term studies have higher proporiton of larger scale studies, fewer small scale studies, more unk


```

#### For each spatial scale, what is the temporal scale?
```{r, echo=FALSE, warning=FALSE}
spat_temp_methods_df %>% 
  group_by(methods_used,num_years,spat_scale) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = spat_scale, y = count)) +
  geom_bar(position = 'fill', stat='identity', aes(fill = num_years)) +
  geom_label(data = spat_temp_methods_df %>% group_by(spat_scale) %>% summarise(count = n()), 
             aes(label = count, y = 1.05)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) +
  ylab('Proportion of studies')


```

#### For each type of study (methods), what is the temporal scale?
Not sure how to group the studies with multiple types of methods here...

```{r, echo=FALSE, warning=FALSE}
spat_temp_methods_df %>%
  group_by(methods_used,num_years,spat_scale) %>%
  summarise(count = n()) %>%
  left_join(spat_temp_methods_df %>% group_by(methods_used) %>% summarise(count = n()), by = "methods_used") %>%
  ggplot(aes(x = fct_reorder(methods_used, -count.y), y = count.x)) +
  geom_bar(position = 'fill', stat='identity', aes(fill = num_years)) +
  geom_label(aes(label = count.y, y = 1.05)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) +
  ylab('Proportion of studies')
```

#### For each temporal scale, what is the type of study (methods)?
Not sure how to group the studies with multiple types of methods here...
```{r, echo=FALSE, warning=FALSE}
spat_temp_methods_df %>%
  group_by(methods_used,num_years,spat_scale) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = num_years, y = count)) +
  geom_bar(position = 'fill', stat='identity', aes(fill = methods_used)) +
  geom_label(data = spat_temp_methods_df %>% group_by(num_years) %>% summarise(count = n()),
             aes(label = count, y = 1.05)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,hjust=1)) +
  ylab('Proportion of studies')
```

